---
output: github_document
---

[![Travis build status](https://travis-ci.org/moodymudskipper/doubt.svg?branch=master)](https://travis-ci.org/moodymudskipper/doubt)

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# doubt

*doubt* modifies the operator `?` to allow definition of new operators (unary, 
binary or n-ary). We refer to those as "dubious" operators, both as a reference to the 
package name and to emphasize the fact that they're not parsed as proper operators.

### Installation

Install with :

``` r
remotes::install_github("moodymudskipper/doubt")
```

### Standard usage

Standard usage as documented in `?utils::Question` still works.

### Call any accessible function as an infix operator

Once the package *doubt* is attached, any accessible function can be called as an
infix operator. See example below:

```{r}
library(doubt, warn.conflicts = FALSE)
1:4 %%intersect? 3:8
```

The precedence is given by the first symbol of the operator, so we can also
do things like:

```{r}
cars *head? 2 +rbind? cars *tail? 2
```

For most cases sticking with `%%fun?` dubious operators will be more intuitive
and readable as we're used to `%fun%` operators and their precedence.

User defined functions will work just the same :

```{r}
AND <- function(x,y) paste(x,"AND",y)
"this" %%AND? "that" 
```

`<-fun?` operators works consistently with the principles outlined above.

```{r}
# in place append
append2 <- function(x, values){
  parent <- parent.env(environment())
  assign(deparse(substitute(x)), append(x, values),envir = parent)
}
x <- 1:2
x <-append2? 11:12 
x
```

It would be convenient however to have the assignment behavior by default 
and be able to do just `x <-append? 11:12` in the case above. This might be 
implemented in near future and will break the above behavior. 

The `?fun?` operators are a special case outlined in the next section.

### `?fun?` n-ary operators

`?fun?` operators are n-ary and have the precedence of `?`, which is right between
`=` (which has the lowest precedence, despite what is documented in `?Syntax`),
and `<-`. 

Any function call `fun(x, y, z)` can be called by `?fun? x ? y ? z`, as in
the following example :

```{r}
# in binary form, works as other dubious ops
cars ?subset? speed == 25
# equivalent to : 
?subset? cars ? speed == 25
# We can add more arguments
cars ?subset? speed == 25 ? "dist"
# equivalent to
?subset? cars ? speed == 25 ? "dist"
```

Note that because of operator precedence `res <- cars ?subset? speed == 25` will
not assign `cars ?subset? speed == 25` to `res`. In that case we can do either
of the following :

```{r}
# Use `=` as it has the lowest precedence
res1 = cars ?subset? speed == 25
res1
res2 = ?subset? cars ? speed == 25
res2
```

Named arguments are not supported at the moment.

Note that because of the very low precedence of these operators and the ambiguity
that wold come from the use of the n-ary notation, only one `?fun?` operator can be
used by statement.

### *piping* with any precedence

We sometimes run into precedence issues when using *magrittr*'s piping operator 
`%>%`. *doubt* wraps this operator and it can be called with the desired precedence,
which is convenient when we want to pipe to a sum for example.

A common use case is the usage of `plotly::ggplotly()` on the result of a
ggplot call. The following will produce an interractive *plotly* chart.

```{r, eval = FALSE}
library(ggplot2)
library(plotly)
ggplot(cars,aes(speed, dist)) + geom_point() +.? ggplotly()
# equivalent to :
ggplotly(ggplot(cars,aes(speed, dist)) + geom_point())
```


### Custom operators

It's generally wiser to define a regular function and rely on the behaviors
described in above sections but one might want to define a `?!` operator
for instance.

Doing so is as simple defining it like any regular function, for instance :

```{r}
`?!` <- function(x) paste(x,"!!")
?! "hello"
```

It will have `?`'s precedence and will allow n-ary notation just like `?fun?`
operators :

```{r}
`?!?` <- function(x,y) paste(x,y,"!!")
?!? "hello" ? "world"
```

Dubious operators can be built from any sequence of characters containing at least one `?` as long as `x <dubious op> y` is syntactic and that the first `?` is binary (it
must be able to "see" the lhs).

The latter statement means that `?!` and `!.?` are allowed but `!?` is not. The
function `arity` can be used to clear confusion, it returns `"invalid"` for
invalid dubious ops, and the arity of the operator otherwise (which is decided
by the first symbol of the operator).

```{r}
arity("?!")
arity("?!?")
arity("!?")
arity("!.?")
```

To be recognized by *doubt* a custom operator must either :

* be defined in the global environment
* be defined in a package AND registered

### Registering a custom dubious op in a package

To use a custom dubious operator your package should import *doubt* and reexport \code{doubt::`?`}. You can do so by adding this code to your package if you use
*roxygen2*:

```{r, eval = FALSE}
#' Modified question mark operator
#'
#' Reexported from package *doubt*
#' @inheritParams doubt::`?`
#' @export
`?` <- doubt::`?`
```

Additionally you should add an `.onAttach` function to your package (usually
done in a `zz.R` file) or edit yours so it updates
*doubt*'s options, as in the following : 

```{r}
.onAttach <- function(libname, pkgname) {
  doubt::register_ops(c("?!", "?!!")) # where `?!` and `?!!` would be your ops
  invisible()
}
```

### *dplyr* extension

You can use a convenient syntax to combine the functionality of *dplyr*
verbs with their suffixed `*_if` / `*_at` variants. 

This won't be documented once the package reaches maturity, because : 

* the implementation is very hacky
* it doesn't reall fit into the scope of the package
* it works only after a pipe, which can be confusing

However I might leave the feature as an Easter egg as I like to use it interactively.

```{r}
library(dplyr)
starwars2 <- starwars[1:2,c(1:3,10:12)]
starwars2
starwars2 %>% select(name, ?is.numeric)
starwars2 %>% mutate(!!!?is.numeric := ~./100, !!!?is.character := toupper)
starwars2 %>% transmute(!!!?is.numeric := ~./100, !!!?is.character)

starwars %>% 
  group_by(gender) %>% 
  summarize(
    # `_if` case
    !!!?is.numeric := mean, 
    # standard case
    name = toString(name),
    # `_at` case
    !!!?vars(starts_with("spec")) := ~toString(unique(.)))
```




<!-- # binary syntax to define special assignment functions -->

<!-- The binary syntax of dubious operators is well suited to design special assignment functions or to write to files or to connections, in the examples below: -->

<!-- * The first function modifies the lhs in place by applying the function in the rhs. -->

<!-- * The second function writes a csv. -->

<!-- ```{r, eval = FALSE} -->
<!-- `<-apply?` <- function(x, fun) { -->
<!--   assign(deparse(substitute(x)), fun(x), envir = parent.frame()) -->
<!--   invisible() -->
<!-- } -->

<!-- x <- "abc" -->
<!-- x <-apply? toupper -->
<!-- x -->
<!-- #> [1] "ABC" -->

<!-- `<-csv?` <- function(file, data) { -->
<!--   write.csv(data, file) -->
<!--   invisible() -->
<!-- } -->
<!-- path <- tempfile(fileext = ".csv")  -->
<!-- path <-csv? head(cars, 3) -->
<!-- read.csv(path) -->
<!-- #>   X speed dist -->
<!-- #> 1 1     4    2 -->
<!-- #> 2 2     4   10 -->
<!-- #> 3 3     7    4 -->
<!-- ``` -->

<!-- ### Unary syntax to execute code in another context -->

<!-- The unary syntax is well suited to define functions that will execute the rhs -->
<!-- in another context, either another environment or even another language. -->

<!-- We propose an example with the *Julia* language, using the package *JuliaCall*. -->

<!-- This should probably live in another package when *doubt* gets more mature, but -->
<!-- for now it serves to showcase the potential of *doubt*. -->

<!-- we defined an operator `?julia>` which allows us to run code in a *Julia* session, -->
<!-- `?julia>` doesn't return any object to R (though it print the output of the -->
<!-- *Julia* console) but its sister function `?!julia>` does. -->

<!-- ```{r} -->
<!-- JuliaCall::julia_setup("C:/Users/afabri/AppData/Local/Julia-1.2.0/bin") -->
<!-- ?julia> "a = 5" -->
<!-- ?!julia> "a * 2" -->
<!-- ``` -->

<!-- Julia's syntax is not so different from R's so most Julia code is syntactic in -->
<!-- R. We offer the possibility to skip the quotes in this case. This way we leverage -->
<!-- syntax highlighting and feel much more as if we're coding in *Julia* directly. -->

<!-- ```{r} -->
<!-- ?julia> 'ceil(3.2)' -->
<!-- ?julia> ceil(3.2) -->
<!-- ``` -->

<!-- For some issues of syntax compatibility some workaround were implemented, -->
<!-- for examples *Julia* assignments, if not quoted, should be done using `:=`. -->

<!-- ```{r} -->
<!-- ?julia> 'x = 42' -->
<!-- ?julia> x := 42 -->
<!-- ``` -->

<!-- Going to the next line should be done by typing `++`, and not `;`. -->

<!-- ```{r, eval = FALSE} -->
<!-- ?julia> "z = begin -->
<!--   x = 1 -->
<!--   y = 2 -->
<!--   x + y -->
<!--   end" -->
<!-- #> 3 -->

<!-- ?julia> z := begin ++ -->
<!--   x := 1 ++ -->
<!--   y := 2 ++ -->
<!--   x + y ++ -->
<!--   end -->
<!-- #> 3 -->
<!-- ``` -->

<!-- And terms separated by spaces in *Julia* can be separated by *?* in R to -->
<!-- keep the expression syntactic. -->

<!-- ```{r, eval = FALSE} -->
<!-- ?julia> 'if 1 < 2 "hello" end' -->
<!-- #> "hello" -->
<!-- ?julia> `if` ? 1 < 2 ? "hello" ? end -->
<!-- #> "hello" -->
<!-- ``` -->
